# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r Load Data, echo=TRUE, eval=TRUE}
data<-read.csv(unz("activity.zip", "activity.csv"),header=TRUE, na.strings = "NA")
data$date<-as.Date(data$date)
```

```{r ggplot2, echo=TRUE, eval=TRUE}
library(ggplot2)
```
## What is mean total number of steps taken per day?
First it is calculated the overall steps made each day, and then, a new dataframe is created with the information.
```{r Mean steps, echo=TRUE, eval=TRUE}
Steps<-with(data,sapply(split(x=steps,f=date),sum))
daysteps<-data.frame(Steps,row.names=as.Date(names(Steps)))
```

With this dataframe we can now see the distribution of the total number of steps taken per day.
```{r histrogram, echo=TRUE, eval=TRUE, fig.width=7, fig.height=6}
plot<-ggplot(daysteps)
plot+geom_histogram(aes(Steps),binwidth=2000)+labs(title="Number of steps per day")
```

We can calculate and store the mean steps made each day and also the median value.
```{r mean, echo=TRUE, eval=TRUE}
mean<-mean(Steps,na.rm=TRUE)
mean
median<-median(Steps,na.rm=TRUE)
median
```
The mean of the distribution is **`r mean`** and the median **`r median`**.

## What is the average daily activity pattern?
We create a new dataframe with the mean number of steps in each interval of 5 minutes during the day and plot the time series.

```{r timeseries, echo=TRUE, eval=TRUE, fig.width=7, fig.height=6}
Interval<-with(data,sapply(split(x=steps,f=interval),mean,na.rm=TRUE))
IntervalSteps<-data.frame(as.integer(as.character(names(Interval))),Interval)
names(IntervalSteps)<-c("Interval","MeanSteps")

plot<-ggplot(IntervalSteps)
plot+geom_line(aes(Interval,MeanSteps))
```

```{r max, echo=TRUE, eval=TRUE}
max<-max(IntervalSteps$MeanSteps,na.rm=TRUE)
max
maxInterval<-IntervalSteps$Interval[IntervalSteps$MeanSteps %in% max]
maxInterval
```
The interval with the maximum value is the **`r maxInterval`**, and the mean number of steps made in those 5 minutes was **`r max`**

## Imputing missing values
We want to know if there are rows with missing values.
```{r NA, echo=TRUE, eval=TRUE}
na<-sum(is.na(data))
```
We find out that there are `r na` rows with missing values. To fill these gaps we can replace the missing value with the rounded mean value of the corresponding interval. We define a function that does the job.
```{r Strategy, echo=TRUE, eval=TRUE}
datafull<-data
fill<-function(steps,interval){
    if(is.na(steps)){
        steps<-round(IntervalSteps$MeanSteps[IntervalSteps$Interval %in% interval])
    }
    steps
}

datafull$steps<-mapply(fill,data$steps,data$interval)
```

The histogram of the data has changed:
```{r histrogram full, echo=TRUE, eval=TRUE, fig.width=7, fig.height=6}
Steps<-with(datafull,sapply(split(x=steps,f=date),sum))
daysteps<-data.frame(Steps,row.names=as.Date(names(Steps)))
plot<-ggplot(daysteps)
plot+geom_histogram(aes(Steps),binwidth=2000)+labs(title="Number of steps per day")
```
The shape of the distribution has changed, now there are more days with similar total steps.

```{r mean full, echo=TRUE, eval=TRUE}
meanfull<-mean(Steps,na.rm=TRUE)
meanfull
medianfull<-median(Steps,na.rm=TRUE)
medianfull
```
The mean of the distribution now is is **`r meanfull`** (vs. the old `r mean`) and the median **`r medianfull`** (vs. the old `r median`). So, the mean is still the same, but the median has decreased 3 steps.

## Are there differences in activity patterns between weekdays and weekends?

We use weekdays() function to classify the measures into weekdays or weekend
```{r weekday, echo=TRUE, eval=TRUE}
datafull$weekday<-weekdays(datafull$date)
datafull$weekday[datafull$weekday %in% c("Saturday","Sunday")]<-"Weekend"
datafull$weekday[datafull$weekday %in% c("Monday","Tuesday","Wednesday","Thursday","Friday")]<-"Weekday"
 datafull$weekday<-as.factor(datafull$weekday)
```

We create again a new dataframe with the mean number of steps in each interval of 5 minutes during the day and plot the time series.
```{r timeseries weekday, echo=TRUE, eval=TRUE, fig.width=7, fig.height=6}
rm(mean) # To avoid R misunderstanding mean as a variable
plot<- ggplot(datafull,aes(interval,steps))
plot+stat_summary(fun.y=mean,geom="line") + facet_grid(weekday~.)
```

We can see that in weekdays there are more steps in early hours and in weekends in late intervals.